---
title: "COVID-19 Cases Over Time"
author: "Auralee Walmer"
date: "November 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Visualizing COVID-19 Cases Worldwide 
### Updated November 27th 2020

This exploration is based on a project from DataCamp.com, found here: <https://projects.datacamp.com/projects/870>. The goal here is to explore the trends in COVID-19 cases over time by country. The data is regularly updated, and I pulled an updated version of it on **November 27th, 2020**. Data accessible here: <https://github.com/RamiKrispin/coronavirus/tree/master/csv>.


```{r datahead, message=FALSE}
# Loading Libraries:
library(readr)
library(knitr)
library(ggplot2)
library(dplyr)

# Read in data
coronavirus <- read.csv("coronavirus.csv")
coronavirus$date <- as.Date(coronavirus$date)
kable(coronavirus[1:5,], caption = "A piece of the data.")

```

## Confirmed Cases Throughout the World

Here's the cumulative trend in cases over time:

```{r pressure, message=FALSE}
# Build data frame consisting of cumulative cases
cumulative <- coronavirus %>% 
  group_by(date, country) %>%
  summarise(value = sum(cases)) %>%
  mutate(new_total_per_day = cumsum(value))
cumulative <- cumulative[cumulative$country=="Zimbabwe",] # final country of the count
cumulative <- cumulative[c("date","new_total_per_day")]
## Cumulative cases over time
cumulative$cumulative_over_time <- NA # prepare column for populating
for (x in 1:nrow(cumulative)) {
  cumulative$cumulative_over_time[x] <- sum(cumulative$new_total_per_day[1:x])
}

# Draw a line plot of cumulative cases vs. date
library(scales)
ggplot(cumulative, aes(x=date, y=cumulative_over_time)) +
geom_line() +
  xlab("Date") + ylab("Cumulative Cases") +
  labs(title = "Cumulative Confirmed Cases Worldwide",
       caption = "Data from the Johns Hopkins University Center for Systems Science and Engineering Coronavirus repository.") +
  scale_x_date(date_labels = "%b %Y", breaks = "2 months") + 
  scale_y_continuous(n.breaks=15, labels = comma)

```

## Comparing China with the Rest of the World

```{r china, message=FALSE, include=FALSE}
china <- coronavirus[coronavirus$country=="China",] ## NOTE THERE ARE REPEAT DATES BC PROVINCE
not_china <- coronavirus[!coronavirus$country=="China",]

## not china
not_china_c <- by(not_china$cases, not_china$date, sum)
not_china_c <- as.data.frame.table(not_china_c)
#not_china_c <- as.data.frame(do.call(rbind,as.list(temp)))
colnames(not_china_c) <- c("date","total cases per day")
## Cumulative cases over time
not_china_c$cumulative_over_time <- NA # prepare column for populating
for (x in 1:nrow(not_china_c)) {
  not_china_c$cumulative_over_time[x] <- sum(not_china_c$`total cases per day`[1:x])
}
not_china_c$is_china <- "Not China"


## just china
china_c <- by(china$cases, china$date, sum)
china_c <- as.data.frame.table(china_c)
colnames(china_c) <- c("date","total cases per day")
china_c$cumulative_over_time <- NA # prepare column for populating
for (x in 1:nrow(china_c)) {
  china_c$cumulative_over_time[x] <- sum(china_c$`total cases per day`[1:x])
}
china_c$is_china <- "China"


## join the data frames
# stack merge
cumulative_comparison <- merge(x=china_c, y=not_china_c, by="is_china", all=TRUE)
#combine columns:
cumulative_comparison$date.x[is.na(cumulative_comparison$date.x)] <- cumulative_comparison$date.y[is.na(cumulative_comparison$date.x)]

cumulative_comparison$cumulative_over_time.x[is.na(cumulative_comparison$cumulative_over_time.x)] <- cumulative_comparison$cumulative_over_time.y[is.na(cumulative_comparison$cumulative_over_time.x)]

cumulative_comparison <- cumulative_comparison[c("is_china","date.x","total cases per day.x", "cumulative_over_time.x")]
colnames(cumulative_comparison) <- c("Is China?","Date","Total New Cases per Day", "Cumulative Cases Over Time")

cumulative_comparison$Date <- as.Date(cumulative_comparison$Date)

```

```{r chinaplot, message=FALSE}
ggplot(cumulative_comparison) +
geom_line(aes(x=Date, y=`Cumulative Cases Over Time`, group=`Is China?`, color=`Is China?`)) +
  xlab("Date") + ylab("Cumulative Cases") +
  labs(title = "Cumulative Confirmed Cases Worldwide",
       caption = "Data from the Johns Hopkins University Center for Systems Science and Engineering Coronavirus repository.") +
  scale_x_date(date_labels = "%b %Y", breaks = "2 months") + 
  scale_y_continuous(n.breaks=15, labels = comma)

```
```{r chinaplot_zoomed, message=FALSE}
ggplot(cumulative_comparison) +
geom_line(aes(x=Date, y=`Cumulative Cases Over Time`, group=`Is China?`, color=`Is China?`)) +
  xlab("Date") + ylab("Cumulative Cases") +
  labs(title = "Cumulative Confirmed Cases Worldwide",
       caption = "Data from the Johns Hopkins University Center for Systems Science and Engineering Coronavirus repository.") +
  scale_x_date(date_labels = "%b %Y", breaks = "1 month", limits=c(as.Date("2020-01-22"),as.Date("2020-05-01"))) + 
  scale_y_continuous(n.breaks=10, labels = comma, limits = c(0,200000))

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
