---
title: "COVID-19 Cases Over Time"
author: "Auralee Walmer"
date: "November 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Visualizing COVID-19 Cases Worldwide 
### Updated November 27th 2020

This exploration is based on a project from DataCamp.com, found here: <https://projects.datacamp.com/projects/870>. The goal here is to explore the trends in COVID-19 cases over time by country. The data is regularly updated, and I pulled an updated version of it on **November 27th, 2020**. Data accessible here: <https://github.com/RamiKrispin/coronavirus/tree/master/csv>.


```{r datahead, message=FALSE}
# Loading Libraries:
library(readr)
library(knitr)
library(ggplot2)
library(dplyr)

# Read in data
coronavirus <- read.csv("/Volumes/Seagate Backup Plus Drive/DOWNLOADS/coronavirus.csv")
kable(coronavirus[1:5,], caption = "A piece of the data.")

```

## Confirmed Cases Throughout the World

Here's the cumulative trend in cases over time:

```{r pressure, message=FALSE}
# Build data frame consisting of cumulative cases
cumulative <- coronavirus %>% 
  group_by(date, country) %>%
  summarise(value = sum(cases)) %>%
  mutate(new_total_per_day = cumsum(value))
cumulative <- cumulative[cumulative$country=="Zimbabwe",] # final country of the count
cumulative <- cumulative[c("date","new_total_per_day")]
cumulative$date <- as.Date(cumulative$date)
cumulative$cumulative_over_time <- NA # prepare column for populating
for (x in 1:nrow(cumulative)) {
  cumulative$cumulative_over_time[x] <- sum(cumulative$new_total_per_day[1:x])
}

# Draw a line plot of cumulative cases vs. date
ggplot(cumulative, aes(x=date, y=cumulative_over_time)) +
geom_line() +
ylab("Cumulative confirmed cases")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
